---
title: "Domaneschi et al. 2017 replication notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(magrittr)
library(haven)
library(lme4)
library(lmerTest)
```

# Preliminaries

Read data

```{r}
df <- read_csv("expt1_cleaned.csv") 
```

from XLS file:
+ Response 1 = Causal Option
+ Response 2 = Other Option
+ Response 3 = Decsriptive Option

```{r}
df %<>% mutate(Response = factor(Response, 
                                 levels = c(1, 2, 3), 
                                 labels = c("Causal","Other","Descriptive")))
```

# Analyses

## Descriptives

> In the Standard condition, our participants chose the causal option 57.77% of the time and the descriptive one 26.66% of the time, while they selected the other option possibility only 15.55% of the time. 

> In the Relevance condition, participants selected the causal inter- pretation 16.66% of the time, and the descriptive one 66.66% of the time, while they chose the other option 16.66% of the time. 


```{r}
df %>%
  group_by(Condition, Response) %>%
  summarise(n = n()) %>%
  mutate(prop = n / sum(n)) %>%
  knitr::kable(digits = 2)
```

**Sucess**

## $\chi^2$ tests.

### Standard condition.

```{r}
ms <- df %>%
  group_by(Condition, Response) %>%
  summarise(n = n()) 

standard <- filter(ms, Condition == "Standard")
```

> The Chi-square statistics confirmed that there was a significant difference between (i) the causal and the descriptive options (v2 (1) = 41.61; p < 0.001); 

```{r}
chisq.test(filter(standard, Response %in% c("Causal","Descriptive"))$n, 
           rescale.p=TRUE)
```

> (ii) the causal and the other option (v2 (1) = 19.96; p < 0.001); 

```{r}
chisq.test(filter(standard, Response %in% c("Causal","Other"))$n, 
           rescale.p=TRUE)
```

> and (iii) the descriptive and the other option (v2 (1) = 4.52; p < 0.05).


```{r}
chisq.test(filter(standard, Response %in% c("Descriptive","Other"))$n, 
           rescale.p=TRUE)
```

### Relevance condition.

```{r}
relevance <- filter(ms, Condition == "Relevance")
```

> The Chi-square statistics confirmed that, while the difference between the causal interpretation and the other option was not significant (v2 (1) = 1.84; p = n.s.), 

```{r}
chisq.test(filter(relevance, Response %in% c("Causal","Other"))$n, 
           rescale.p=TRUE)
```

**Major discrepancy**: they report $\chi^2 = 1.84$ but by their own descriptives, these are literally identical measurements, so how could they have $\chi^2 > 0$?

> the percentage of choice for the descriptive option was significantly higher than that of the causal option (v2 (1) = 27.7; p < 0.001) 


```{r}
chisq.test(filter(relevance, Response %in% c("Descriptive","Causal"))$n, 
           rescale.p=TRUE)
```

> and that of the other option preference (v2 (1) = 27.7; p < 0.001).


```{r}
chisq.test(filter(relevance, Response %in% c("Descriptive","Other"))$n, 
           rescale.p=TRUE)
```

## Across conditions.

> The above trend of results was further confirmed in the across- conditions comparison. Statistical analysis was performed with Linear Mixed Models (LMM) using the lme4 package (Bates et al., 2015) and the lmerTest package to extend the lmer model and pro- vide the degrees of freedom and the p-value. The random structure of the model was constructed by using subjects and items as ran- dom factors (Baayen, Davidson, & Bates, 2008), while participants’ responses in the two conditions (i.e., Standard vs. Relevance) con- stituted the fixed structure of the model. The results revealed a sig- nificant effect of condition (t = 2.81, df = 1, p < 0.01), further confirming the reversed pattern of choices between the causal and descriptive option in the two conditions (see Fig. 2). 

Problem: Response is actually a nominal variable with three different options, not clear how this has been coded in the LMM.

Let's try putting these in as a factor and see what comes out. 

```{r}
mod <- lmer(Response ~ Condition  + (1|`Subject_ID`) + (1|`Item_ID`), 
     data = df)

summary(mod)
```

```{r}
mod <- lmer(as.numeric(Response) ~ Condition  + (1|Subject_ID) + (1|Item_ID), 
     data = df)

summary(mod)
```
